---
title: "Lab06 - Transformations for ANOVA"
output: pdf_document
---

## Goals

The goal in this lab is to practice working with transformations for ANOVA.

## Loading packages

Here are some packages with functionality you may need for this lab.  Run this code chunk now.

```{r}
library(readr)
library(ggplot2)
library(gridExtra)
library(mosaic)
library(dplyr)

options("pillar.sigfig" = 10) # print 10 significant digits in summarize output
```

A gas chromatograph is an instrument that measures the amounts of various compounds contained in a sample by separating the various constituents.  The total number of counts recorded by the chromatograph is proportional to the amount of the compound present.

A calibration experiment was performed to see how the recorded counts from the chromatograph related to the concentration of a compound in a mixture and the flow rate through the chromatograph.  In this lab we will just look at the relationship between the concentration (explanatory variable) and the counts (response variable).

```{r}
chromatography <- read_csv("http://www.evanlray.com/data/sdm3/Chapter_29/Ch29_Chromatography.csv")
names(chromatography) <- c("concentration", "flow_rate", "counts")

chromatography %>%
  count(concentration)
```

#### 1. Make an appropriate plot of the data: it might be nice to use a histogram or density plot, separately for each value of cylinders. Also calculate the standard deviation for each group. Would it be appropriate to use an ANOVA model for these data?

```{r}
ggplot(data = chromatography, mapping = aes(x = concentration, y = counts)) +
  geom_boxplot()

chromatography %>%
  group_by(concentration) %>%
  summarize(
    sd(counts)
  )
```

#### 2. Find a transformation of the data so that the ANOVA model would be appropriate.

```{r}
chromatography <- chromatography %>%
  mutate(
    sqrt_counts = sqrt(counts),
    log_counts = log(counts),
    inv_sqrt_counts = -1/sqrt(counts)
  )

p_sqrt <- ggplot(data = chromatography, mapping = aes(x = concentration, y = sqrt_counts)) +
  geom_boxplot() 
p_log <- ggplot(data = chromatography, mapping = aes(x = concentration, y = log_counts)) +
  geom_boxplot()
p_inv_sqrt <- ggplot(data = chromatography, mapping = aes(x = concentration, y = inv_sqrt_counts)) +
  geom_boxplot()

grid.arrange(p_sqrt, p_log, p_inv_sqrt)

chromatography %>%
  group_by(concentration) %>%
  summarize(
    sd(sqrt_counts),
    sd(log_counts),
    sd(inv_sqrt_counts)
  )
```

The standard deviations in the different groups are most consistent with a log transformations.  The distributions are close enough to normally distributed.

#### 3. Conduct a test of the claim that the mean count is the same for all three concentration levels.

This is formally a test of the claim that the mean of the log count is the same for all three concentration levels, since we're working with log-transformed data.

$H_0: \mu_1 = \mu_2 = \mu_3$
$H_A:$ at least one of $\mu_1$, $\mu_2$, and $\mu_3$ is not equal to the others

```{r}
model_fit <- lm(log_counts ~ concentration, data = chromatography)
summary(model_fit)
```

The p-value for the test is less than $2.2 \times 10^{-16}$.  There is extremely strong evidence against the null hypothesis that the mean of the log count is the same for all three concentration levels.

#### 4. Report and interpret an estimate of the difference in the centers of the distributions of counts for the high concentration and the low concentration, as well as a 95% confidence interval for that difference.  You should be able to do this in a few different ways.

```{r}
fit <- lm(log_counts ~ concentration, data = chromatography)
summary(fit)
confint(fit)
```


```{r}
exp(-2.31775)
```
We estimate that the median count in the low group is about 0.1 times the median count for the high group.

We can verify this by looking at the actual medians for these groups:

```{r}
chromatography %>%
  group_by(concentration) %>%
  summarize(
    median(counts)
  )
```

OR

```{r}
exp(2.31775)
```
We estimate that the median count in the high group is about 10 times the median concentration for the low group.
